---
title: "CSE 160 - Final Project - Chicago COVID-19 Risk"
author: "Gabby Rub, Jacklyn Clauss, Mario Martino, Shuang Lin, Tony Wu"
date: "22 November 2022"
output: html_notebook
---

Import data and Chicago shapefile
```{r}

df_covid <- read.csv("COVID-19_Cases__Tests__and_Deaths_by_ZIP_Code.csv")

```

Clean data
```{r}


```

Clean shapefile
```{r}


```

Partition data (80:20)
```{r}


```

Build Linear Regression model. Save performance measures or later comparison.
```{r}



```

Build Naive Bayes model. Save performance measures or later comparison.
```{r}


```

Build Decision Tree model. Save performance measures or later comparison.
```{r}


```

Build K Nearest Neighbor model. Save performance measures or later comparison.
```{r}


```

Compare all model performances (Table & ROC Curves)
```{r}


```

Use best model to predict infection rate on test data.
```{r}


```

Add binary attribute to data frame to create the easier read mapping
```{r}


```

Mapping of infection risk percentages
```{r}


```

Mapping of binary risk areas
```{r}


```
